The Effect of Racial and Ethnic Attitudes on Asian Identity in the U.S

### 2024-09-19 my_paper/manuscript_main.tex
write more

I discuss a conceptual framework of identity in the spirit of \textcite{akerlofEconomicsIdentity2000}. A person belongs to some ethnic group, and their actions either affirm or deny their ethnic identity. Actions that deviate from what is proscribed of the ethnic identity are costly.

Formally, a person $i$ belongs to ethnic group $e_i \in \{H, NH\}$, where $H$ is Asian and $NH$ is non-Asian. Agent $i$'s utility depends on their actions and the extent to which their actions affirm their identity $I_i$:

@@

A person's identity, $I_i$, is influenced by their own actions, the actions of others, and the behavior proscribed by their ethnicity. I write this as:

@@

A person $i$ chooses action $a_i$ that maximizes their utility function given ethnic group $e_i$, proscribed appropriate behavior $\pmb{B}_{e_{i}}$, and the actions of others $\pmb{a_{-i}}$. This implies the following first-order condition (F.O.C.):

@@

Whose solution $a_{i}^{\star}$ yields utility $U_{i}^{\star}$. Now, suppose a person can choose their ethnic identity at a cost of $c$. They will do so if $\tilde{U_{i}}^{\star} \geq  U_{i}^{\star} + c$. Where $\tilde{U_{i}}^{\star}$ is the utility obtained from optimal actions $\tilde{a_i^{\star}}$ under the counterfactual ethnicity.

That is $i$ will change identities when the benefits of doing so $\tilde{U_{i}}^{\star} - U_{i}^{\star}$ exceed the costs $c$. These net benefits are non-zero only if $\frac{d I_i}{d a_i} \neq 0$ and $\frac{\partial U_i}{\partial I_i} \neq 0$. This suggests that an empirical analysis of the determinants of identity choice should focus on: (1) individual characteristics that would lead to different $a_i$ under different identities, (2) contextual characteristics that would lead to different $a_{-i}$---bias---under different identities, (3) the analysis should focus on a sample of the population with small $c$, and (4) the sub-sample with a utility that is greatly affected by their identity---i.e., $\frac{\partial U_i}{\partial I_i} \neq 0$). From the empirical analysis, I could investigate the characteristics that would affect $i$'s actions to take different identities from point (1). These characteristics could be the generation immigrants belong to, whether their parents are interethnic or endogamous, etc. I also investigate how different state-level biases could affect identity. Finally, restricting the sample to people with a small cost of changing identity $c$ guarantees that I do not include populations that would never change identities otherwise---for example, non-Asian Whites with non-Asian ancestry.

@@

In this section, I describe the datasets I use. To study the association between social attitudes and self-reported Asian identity, I must measure subjective and objective Asian identities to select a subgroup of Asian immigrants for analysis. Thus, I use the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) \autocite{floodsarahIntegratedPublicUse2021a} and use information on ancestry to construct an objective identity measure. I also use implicit association test (IAT) as a measure of bias, in which bias will shift the costs of identifying as a particular ethnic identity (Asian).

@@

The objective measure of identity---unlike the self-reported measure where respondents answer affirmatively when asked if they are Asian or Latino---depends on the birthplaces of the individual, their two parents, and four grandparents. Thus, the three identifiable generations are: 1) first-generation immigrants that are born in a Asian country with both parents also being born in a Asian country, 2) second-generation immigrants are native-born citizens to at least one parent that was born in a Asian country, 3) third-generation immigrants are native-born citizens to two native-born parents and at least one grandparent that was born in a Asian country.\footnote{I restrict first-generation immigrants whose parents were born in a Spanish country to avoid including naturally born US citizens that were born abroad to US parents.} I restrict the sample to Asian Whites, first-, second-, and third-generation immigrants who are 17 year old and younger and still live with their parents between 2004 and 2021. I present a summary of the sample statistics in the Table (\ref{tab:sumstat1}).

The overall sample is 49\% female, and 91\% of the sample self-report their identity as Asian---answered yes to the question ``are you Latino/Asian?''. The average age is 8.6-year-old. Almost 14\% of mothers have a college degree, and 14\% of fathers have a college degree. I provide the rest of the summary statistics for the overall sample and for both the overall sample and each generation in Table (\ref{tab:sumstat1}).

Moreover, using the place of birth of parents and grandparents, I can objectively identify their ethnic ancestry. Consequently, I can identify different types of parents and grandparents. Using the place of birth of parents, I can divide parents of second-generation children into three objective types: 
\begin{enumerate}
\item Objectively Asian-father-Asian-mother (HH)
\item Objectively Asian-father-White-mother (HW)
\item Objectively White-father-Asian-mother (WH)
\end{enumerate}

Similarly, using the place of birth of grandparents, I can divide grandparents of third-generation children into 15 objective types: (1) objectively Asian paternal grandfather-Asian paternal grandmother-Asian maternal grandfather-Asian maternal grandmother (HHHH); (2) objectively White paternal grandfather-Asian paternal grandmother-Asian maternal grandfather-Asian maternal grandmother (WHHH); (3) objectively Asian paternal grandfather-White paternal grandmother-Asian maternal grandfather-Asian maternal grandmother (HWHH), etc...

My analysis uses a sub-sample of the US population; I show in Table (\ref{tab:hispbygen}) that I have enough observations in each generation. Consistent with the literature on ethnic attrition among Asians, I find significant attrition among third-generation Asian immigrants.\footnote{In \textcite{duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018, antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}, the authors find substantial attrition among Asians.} These results are displayed in Table (\ref{tab:hispbygen}): most first- and second-generation Asian immigrants self-report their identity as Asian. Of the first-generation Asian immigrants, 96\% self-report their identity as Asian. Similarly, 95\% of the second-generation Asian immigrants self-report their identity as Asian, and 85\% of third-generation Asian immigrants identified as Asian. That is a more than threefold increase in attrition rates. Most of the attrition among third-generation Asians is driven by attrition among the children of interethnic marriages.

@@

I use skin tone implicit association test data to construct a proxy of state-level prejudice \autocite{greenwaldMeasuringIndividualDifferences1998}. This measure has been used in the social sciences, especially in psychology. Previous work has shown that IAT test scores are hard to manipulate \autocite{egloffPredictiveValidityImplicit2002}.

@@

I construct another proxy measure of racial animus using the ANES survey \autocite{anes2021} to measure animus, or discrimination, against Black Americans. ANES is a survey that has been conducted since 1948 and is widely used in political science. The survey asks respondents about their attitudes toward different racial groups, voting intentions, and other political questions. I use several questions from the ANES surveys conducted between 2004 and 2020 to measure racial animus. The racial animus index is constructed by taking the average of the responses to several questions measuring racial animus. \footnote{The questions used are similar to those used by \textcite{charlesPrejudiceWagesEmpirical2008}. The questions are: (1) ``Conditions Make it Difficult for Blacks to Succeed'', (2) ``Blacks Should Not Have Special Favors to Succeed'', (3) ``Blacks Must Try Harder to Succeed'', (4) ``Blacks Gotten Less than They Deserve Over the Past Few Years'', and (5) ``Feeling Thermometer Toward Asians.''}

@@

\begin{align}
H_{ist}^g &= \beta_1^g Bias_{st} + \beta_2^g DadCollegeGrad_{ist} + \beta_3^g MomCollegeGrad_{ist} \nonumber \\ 
            &+ \beta_4^g Women_{ist} + X_{ist}^g\pi + \gamma_{rt} 
           + \varepsilon_{ist}; 
           \text{where } g \in \{1,2,3\} \label{eq:identity_reg_bias}
\end{align}

Where $H_{ist}^g$ be the self-reported Asian identity of person $i$ in state $s$ at the time of interview $t$, let $Bias_{st}$ be the average state-level bias in state $s$ at time $t$, $DadCollegeGrad_{ist}$, and $MomCollegeGrad_{ist}$ are indicator variables that are equal to one if the father or mother graduated from college, $Women_{ist}$ is an indicator variable for sex, and $X_{ist}$ is a vector of controls.\footnote{The controls include quartic age, fraction of population that is Asian in state $s$, type of parents (WH, HW, or HH), type of grandparents (HHHH, HHHW, etc.), and dummy variables the generation to which person $i$ belong.} Additionally, $\gamma_{rt}$ is region-time fixed effects that controls for region $\times$ year specific shocks.\footnote{I do not include state fixed effects because of lack of with-in state variation in bias.} The region $\times$ year also controls for systematic differences between regions in the overall Asian population and bias toward Asians, even if they vary over time. Throughout the analysis, I cluster the standard errors at the state level to account for correlation in the error term $\varepsilon_{ist}$ within a state, overtime.

@@

I report the main results of estimating equation (\ref{eq:identity_reg_bias}) in Figure (\ref{plot01-regression-gen}). I present the results of estimating the main specification for second-generation immigrants in panel (A) and for sub-samples of HH, HW, and WH children in panels (B), (C), and (D), respectively. I find that bias and self-reported Asian identity are negatively associated, while parental education and self-reported Asian identity are positively associated. A one standard deviation increase in state-level bias is associated with a 4 percentage points decrease in self-reported Asian identity. Among first- and second-generation Asian immigrants, a one standard deviation increase in state-level bias is associated with 2 and 4 percentage points decrease in self-reported Asian identity. The coefficient is not statistically significant for first-generation, but the confidence interval is mostly in the negative territory.

I report the results of the same regression but on sub-samples of second-generation immigrants by type of parents---interethnic and endogamous parents---in Figure (\ref{plot01-regression-byparent}). I present the results of estimating the main specification on second-generation immigrants in panel (A) and on sub-samples of HH, HW, and WH children in panels (B), (C), and (D), respectively. I find that children of endogamous Asian marriages are more influenced by state-level bias than those of interethnic marriages. I find that a one standard deviation increase in state-level bias is associated with 5 percentage points decrease in self-reported Asian identity among children of HH parents.

>>>>

I discuss a conceptual framework of identity in the spirit of \textcite{akerlofEconomicsIdentity2000}. A person belongs to some racial group, and their actions either affirm or deny their racial identity. Actions that deviate from what is proscribed of the racial identity are costly.

Formally, a person $i$ belongs to racial group $e_i \in \{A, W\}$, where $A$ is Asian and $W$ is White. Agent $i$'s utility depends on their actions and the extent to which their actions affirm their identity $I_i$:

@@

A person's identity, $I_i$, is influenced by their own actions, the actions of others, and the behavior proscribed by their race. I write this as:

@@

A person $i$ chooses action $a_i$ that maximizes their utility function given racial group $e_i$, proscribed appropriate behavior $\pmb{B}_{e_{i}}$, and the actions of others $\pmb{a_{-i}}$. This implies the following first-order condition (F.O.C.):

@@

Whose solution $a_{i}^{\star}$ yields utility $U_{i}^{\star}$. Now, suppose a person can choose their racial identity at a cost of $c$. They will do so if $\tilde{U_{i}}^{\star} \geq  U_{i}^{\star} + c$. Where $\tilde{U_{i}}^{\star}$ is the utility obtained from optimal actions $\tilde{a_i^{\star}}$ under the counterfactual race.

That is $i$ will change identities when the benefits of doing so $\tilde{U_{i}}^{\star} - U_{i}^{\star}$ exceed the costs $c$. These net benefits are non-zero only if $\frac{d I_i}{d a_i} \neq 0$ and $\frac{\partial U_i}{\partial I_i} \neq 0$. This suggests that an empirical analysis of the determinants of identity choice should focus on: (1) individual characteristics that would lead to different $a_i$ under different identities, (2) contextual characteristics that would lead to different $a_{-i}$---bias---under different identities, (3) the analysis should focus on a sample of the population with small $c$, and (4) the sub-sample with a utility that is greatly affected by their identity---i.e., $\frac{\partial U_i}{\partial I_i} \neq 0$). From the empirical analysis, I could investigate the characteristics that would affect $i$'s actions to take different identities from point (1). These characteristics could be the generation immigrants belong to, whether their parents are interracial or endogamous, etc. I also investigate how different state-level biases could affect identity. Finally, restricting the sample to people with a small cost of changing identity $c$ guarantees that I do not include populations that would never change identities otherwise---for example, non-Asian Whites with non-Asian ancestry.

@@

In this section, I describe the datasets I use. To study the association between social attitudes and self-reported Asian identity, I must measure subjective and objective Asian identities to select a subgroup of Asian immigrants for analysis. Thus, I use the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) \autocite{floodsarahIntegratedPublicUse2021a} and use information on ancestry to construct an objective identity measure. I construct a composite measure of bias using the implicit association test, the American National Election Studies, and hate crimes against Asians. My composite is created using the method of \textcite{lubotskyInterpretationRegressionsMultiple2006} to reduce attenuation bias.

@@

The objective measure of identity---unlike the self-reported measure where respondents answer affirmatively when asked if they are Asian---depends on the birthplaces of the individual, their two parents, and four grandparents. Thus, the three identifiable generations are: 1) first-generation immigrants that are born in an Asian country with both parents also being born in an Asian country, 2) second-generation immigrants are native-born citizens to at least one parent that was born in an Asian country, 3) third-generation immigrants are native-born citizens to two native-born parents and at least one grandparent that was born in an Asian country.\footnote{I restrict first-generation immigrants whose parents were born in a Spanish country to avoid including naturally born US citizens that were born abroad to US parents.} I restrict the sample to Asian, first-, second-, and third-generation immigrants who are 17 year old and younger and still live with their parents between 2004 and 2021. I present a summary of the sample statistics in the Table (\ref{tab:sumstat1}).

The overall sample is 49\% female, and 65\% of the sample self-report their identity as Asian---answered yes to the question ``what is your race''. The average age is 8.4-year-old. Almost 52\% of mothers have a college degree, and 52\% of fathers have a college degree. I provide the rest of the summary statistics for the overall sample and for both the overall sample and each generation in Table (\ref{tab:sumstat1}).

Moreover, using the place of birth of parents and grandparents, I can objectively identify their ethnic ancestry. Consequently, I can identify different types of parents and grandparents. Using the place of birth of parents, I can divide parents of second-generation children into three objective types: 
\begin{enumerate}
\item Objectively Asian-father-Asian-mother (AA)
\item Objectively Asian-father-White-mother (AW)
\item Objectively White-father-Asian-mother (WA)
\end{enumerate}

Similarly, using the place of birth of grandparents, I can divide grandparents of third-generation children into 15 objective types: (1) objectively Asian paternal grandfather-Asian paternal grandmother-Asian maternal grandfather-Asian maternal grandmother (AAAA); (2) objectively White paternal grandfather-Asian paternal grandmother-Asian maternal grandfather-Asian maternal grandmother (WAAA); (3) objectively Asian paternal grandfather-White paternal grandmother-Asian maternal grandfather-Asian maternal grandmother (AWAA), etc...

My analysis uses a sub-sample of the US population; I show in Table (\ref{tab:hispbygen}) that I have enough observations in each generation. Consistent with the literature on ethnic attrition among Asians, I find significant attrition among third-generation Asian immigrants.\footnote{In \textcite{duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018, antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}, the authors find substantial attrition among Asians.} These results are displayed in Table (\ref{tab:hispbygen}): most first- and second-generation Asian immigrants self-report their identity as Asian. Of the first-generation Asian immigrants, 96\% self-report their identity as Asian. 73\% of the second-generation Asian immigrants self-report their identity as Asian, and 31\% of third-generation Asian immigrants identified as Asian. The attrition among second- and third-generation Asian immigrants is primarily driven by children of interracially married parents.

@@

I use Asian implicit association test data to construct a proxy of state-level prejudice \autocite{greenwaldMeasuringIndividualDifferences1998}. This measure has been used in the social sciences, especially in psychology. Previous work has shown that IAT test scores are hard to manipulate \autocite{egloffPredictiveValidityImplicit2002}.

@@

I construct another proxy measure of racial animus using the ANES survey \textcite{anes2021} to measure animus, or discrimination, against Black Americans. ANES is a survey that has been conducted since 1948 and is widely used in political science. The survey asks respondents about their attitudes toward different racial groups, voting intentions, and other political questions. I use several questions from the ANES surveys conducted between 2004 and 2020 to measure racial animus. The racial animus index is constructed by taking the average of the responses to several questions measuring racial animus. \footnote{The questions used are similar to those used by \textcite{charlesPrejudiceWagesEmpirical2008}. The questions are: (1) ``Conditions Make it Difficult for Blacks to Succeed'', (2) ``Blacks Should Not Have Special Favors to Succeed'', (3) ``Blacks Must Try Harder to Succeed'', (4) ``Blacks Gotten Less than They Deserve Over the Past Few Years'', and (5) ``Feeling Thermometer Toward Asians.''}

Lastly, I incorporate data from the Uniform Crime Reports (UCR) \autocite{ucrbook} to quantify state-level hate crimes against Asians. Hate crime data offers a tangible measure of racially motivated aggression and discrimination, which, when combined with implicit and explicit bias measures, allows for a fuller understanding of the landscape of prejudice across states. This combination of implicit and explicit bias measures, along with hate crime statistics, offers a multidimensional approach to understanding the nature and prevalence of racial prejudice.

@@

\begin{align}
A_{ist}^g &= \beta_1^g Bias_{st} + \beta_2^g DadCollegeGrad_{ist} + \beta_3^g MomCollegeGrad_{ist} \nonumber \\ 
            &+ \beta_4^g Women_{ist} + X_{ist}^g\pi + \gamma_{rt} 
           + \varepsilon_{ist}; 
           \text{where } g \in \{1,2,3\} \label{eq:identity_reg_bias}
\end{align}

Where $A_{ist}^g$ be the self-reported Asian identity of person $i$ in state $s$ at the time of interview $t$, let $Bias_{st}$ be the average state-level bias in state $s$ at time $t$, $DadCollegeGrad_{ist}$, and $MomCollegeGrad_{ist}$ are indicator variables that are equal to one if the father or mother graduated from college, $Women_{ist}$ is an indicator variable for sex, and $X_{ist}$ is a vector of controls.\footnote{The controls include quartic age, fraction of population that is Asian in state $s$, type of parents (WA, AW, or AA), type of grandparents (AAAA, AAAW, etc.), and dummy variables the generation to which person $i$ belong.} Additionally, $\gamma_{rt}$ is region-time fixed effects that controls for region $\times$ year specific shocks.\footnote{I do not include state fixed effects because of lack of with-in state variation in bias.} The region $\times$ year also controls for systematic differences between regions in the overall Asian population and bias toward Asians, even if they vary over time. Throughout the analysis, I cluster the standard errors at the state level to account for correlation in the error term $\varepsilon_{ist}$ within a state, overtime.

@@

I report the main results of estimating equation (\ref{eq:identity_reg_bias}) in Figure (\ref{plot01-regression-gen}). I present the results of estimating the main specification for second-generation immigrants in panel (A) and for sub-samples of AA, AW, and WA children in panels (B), (C), and (D), respectively. I find that bias and self-reported Asian identity are negatively associated, while parental education and self-reported Asian identity are positively associated. A one standard deviation increase in state-level bias is associated with a 4 percentage points decrease in self-reported Asian identity. Among first- and second-generation Asian immigrants, a one standard deviation increase in state-level bias is associated with 2 and 4 percentage points decrease in self-reported Asian identity. The coefficient is not statistically significant for first-generation, but the confidence interval is mostly in the negative territory.

I report the results of the same regression but on sub-samples of second-generation immigrants by type of parents---interethnic and endogamous parents---in Figure (\ref{plot01-regression-byparent}). I present the results of estimating the main specification on second-generation immigrants in panel (A) and on sub-samples of AA, AW, and WA children in panels (B), (C), and (D), respectively. I find that children of endogamous Asian marriages are more influenced by state-level bias than those of interethnic marriages. I find that a one standard deviation increase in state-level bias is associated with 5 percentage points decrease in self-reported Asian identity among children of AA parents.
